---
title: "Weighted Matching: estimation and inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weighted Matching: estimation and inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(WeightedMatching)
```

`WeightedMatching` implements fixed-count matching with supplied positive probability
weights. The original self-normalized donor share is
`W_j / sum(W_selected)`. Each query has exactly `M` distinct opposite-arm
donors, with reuse permitted across queries. PATE uses the all-row weighted
target; PATT uses the treated weighted target and only the control imputation.

This tutorial is for the independent `WeightedMatching` package. The original
WDSM package, `wdsmatch` 0.2.1, installs separately; its `wdsmatchATE()` and
`wdsmatchATT()` interfaces are not exported by this package.

This tutorial runs small synthetic interface examples. It is not a coverage
experiment or a verification of the assumptions needed for causal
identification, residual centering, geometry or fitted-score inference. Weights
are known functions of treatment and covariates; they are not estimated or
refitted. No dependent-survey cluster/stratum inference is supplied.

## Prepare rows and weights

```{r data}
set.seed(928)
n <- 400L
X <- cbind(x1 = runif(n, -1, 1), x2 = runif(n, -1, 1))
Z <- rbinom(n, 1, plogis(-0.2 + 0.5 * X[, "x1"] - 0.3 * X[, "x2"]))
W <- exp(0.3 * Z + 0.2 * X[, "x1"])
Y <- 1 + X[, "x1"] + 0.4 * X[, "x2"] +
  Z * (0.8 + X[, "x2"]) + runif(n, -0.5, 0.5)
```

All inputs keep the same original rows. Multiplying every W by a common
positive constant leaves the target ratios and self-normalized donor shares
unchanged. The matching engine uses Euclidean coordinates as supplied; choose
and retain the intended scaling explicitly. It does not duplicate rows using W.

## Supplied coordinates and fitted correction

```{r fixed-map}
fixed_ate <- wm_fit(Y, Z, W, scores0 = X, M = 3,
                    estimand = "PATE", regression = "polynomial", degree = 2)
fixed_att <- wm_fit(Y, Z, W, scores0 = X, M = 3,
                    estimand = "PATT", regression = "polynomial", degree = 2)
c(PATE = fixed_ate$estimate, PATT = fixed_att$estimate)
summary(fixed_ate)$donor_usage
```

`wm_fit()` uses a complete weighted polynomial or the specified tensor spline
in each arm's own matching variables. For supplied predictions, use `wm_match()`
instead. `scores1` can specify a different PATE matching map; PATT needs only
`scores0`. Both APIs retain the graph, donor shares and weighted incoming reuse.
Basis size/allocation guards remain, and the positive-degree polynomial builder
supports up to 100 coordinates. A supplied-map theoretical result does not
remove these computational limits.

The ordinary row variance/contribution calculation requires its documented
strong graph-field residual centering and feasible-correction conditions. A
distinct-map reciprocal inference calculation has a separate bounded contract.
An available point or SE does not establish those premises.

## Fitted arm-specific double scores

Declare the intended intercept and covariate columns explicitly:

```{r fitted-dsm}
D <- cbind(intercept = 1, X)
dsm <- wm_wdsm_fit(Y, Z, W, ps_design = D, pg0_design = D,
                   pg1_design = D, M = 3, estimand = "PATE",
                   inference = "full_x")
dsm_att <- wm_wdsm_fit(Y, Z, W, ps_design = D, pg0_design = D,
                       M = 3, estimand = "PATT", inference = "full_x")
c(PATE = dsm$estimate, PATT = dsm_att$estimate)
```

The wrapper fits a shared probability-weighted PS, separate ordinary arm PG
models, pooled score centers/scales and supplied-W complete quadratic
correction. Its matching maps are `(PS, PG0)` and `(PS, PG1)`; PATT omits unused
treated prediction blocks. Set `ps_weighting="unit"` for the original unweighted
PS-fitting convention while retaining W in the estimator and correction.

`inference="none"` is the point-only default. The explicit `"full_x"` request
includes the full joint PS/PG/scaling/correction influence and covariance,
conditional on correct full-X predictions and the applicable actual-root,
current-graph, moment and numerical-transfer conditions. A correct PS alone
does not establish this route. Finite model lists can be fitted with
`wm_model_fit()`; their matching dimension and complete nuisance dimension
are different. Raw scalar PSM uses a separate documented `scalar_psm` route,
not the corrected double-score statistic.

## Contribution replication

```{r contribution}
contribution <- wm_bootstrap(dsm, B = 99, seed = 17,
                              method = "contribution", conf.level = 0.95)
contribution$conf.int
```

This public fitted-wrapper call binds the retained complete inference inputs
and applies the selected contribution operator. It does not refit nuisance
models or rebuild the matching graph. Ordinary Gaussian contribution intervals
use their analytic conditional variance; B controls generated draws, not that
analytic variance. Other covariance scopes can be prepared explicitly with
`wm_model_inference()` or `wm_fitted_inference()`. Critical generally
non-Gaussian routes have different interval options and are not selected by
this example.

## Original fixed-reuse count refitting

Generate the counts once and share them across intended comparisons:

```{r refit}
set.seed(29)
counts <- rmultinom(49, n, rep(1 / n, n))
refit_ate <- wm_bootstrap(dsm, method = "fixed_reuse", counts = counts)
refit_att <- wm_bootstrap(dsm_att, method = "fixed_reuse", counts = counts)
rbind(PATE = refit_ate$conf.int, PATT = refit_att$conf.int)
refit_ate$variance_divisor
```

For default quadratic fitted wrappers, this call refits the complete retained
PS/PG/scaling/correction pipeline using the same full-n multiplicities in every
active block. W, the original donors and incoming reuse remain fixed. Replicate
PATE/PATT denominators use the corresponding counted target weights. The
original variance convention is divisor B, and intervals are centered on the
original point. Contribution and refit variances need not agree in finite data.

For a nested `wm_match` object, `refit=NULL` holds predictions fixed. Supply a
callback returning `mean0` and, for PATE, `mean1` to implement the intended
complete refit or a separately proved compatible prediction-only procedure.
Nondefault local-polynomial/QR corrections require such an explicit callback;
they are not silently replaced by quadratic fits.

Every prescribed column must succeed. A support/fit failure stops the
calculation; do not delete the column, fill it with zero or redraw it. Fixed-B
replication has Monte Carlo variation. Sampling validity and a growing-B result
need their stated refit, agreement, moment and all-draw numerical conditions.

## Options and reproducibility

Use `?wm_match`, `?wm_fit`, `?wm_wdsm_fit`, `?wm_model_fit`,
`?wm_bootstrap` and `?wm_bootstrap_refit` for argument and return contracts.
Weights remain supplied and known in the current framework. The logistic
fitters reject probabilities outside their fixed numerical acceptance interval;
that finite-data check does not establish asymptotic numerical availability
under an unbounded propensity index.

Record `sessionInfo()`, the actual package/source commit, target, M, maps,
scaling, models, count stream and failed draws. The source repository supplies
separate principal-study, WDSM and Lenis recipes and saved-result replay.
Those workflows are excluded from the installed package. Reproducing a frozen
producer requires its specified checkout rather than silently updating its
source hashes to the latest package.
